\textbf{Discovery of a disrupting open cluster in the Galactic halo:\\ a recent star formation event in the leading arm of the Magellanic stream}

### 2018-11-03 ms.tex
more references

We use astrometric data from the \gaia\ mission (\citealt{Prusti:2016}), data release 2 (\DR{2}; \citealt{Gaia-Collaboration:2018, Lindegren:2018}) to search for distant, comoving multiplets of blue stars.
Our original intent was to search for small, distant, comoving groups of blue horizontal branch stars to identify new candidate satellites of the Milky Way.
We therefore initially select all stars from \gaia\ with parallax $\varpi < 1$, color $-0.5 < (\bprp) < 0$, $G$-band magnitude $G < 20$, and Galactic latitude $|b| > 20^\circ$ (see Appendix~\ref{sec:queries} for database query).
We further exclude stars within a $15^\circ$ radius from the LMC, and a $8^\circ$ radius from the SMC --- 27,895 stars remain after these cuts.
We then cross-match this catalog to itself with both sky positions and proper motions: we search for pairs of stars that have separations $s < 0.5^\circ$ and proper motion differences $|\Delta \mu| < 0.5~\masyr$.
We then combine mutually-connected comoving pairs into small groups of stars that are colocated on the sky and comoving in proper motions, and remove groups that have $<4$ members.
We cross-match the mean sky positions of the groups to locations of local group galaxies \citep{McConnachie:2012} and Milky Way globular clusters \citep[2010 edition;][]{Harris} and filter out all groups that lie within 1 degree of these known objects.
After these filters, just one group of blue, comoving stars remains at $(\textrm{RA}, \textrm{Dec}) \sim (179, -29)^\circ$.

@@

In detail, we start by using the \texttt{isochrones} package \citep{Morton:XXXX} to generate posterior samplings over stellar parameters for each individual source given its photometry and an interim prior.
We use the \acronym{MIST} \citep{Choi:2016} isochrone library, and \texttt{isochrones} automatically performs interpolation between the provided grid of stellar isochrones to evaluate predicted stellar parameters at a given location in the CMD.
The stellar parameters for each source are its ``equal evolutionary point'' number $\eep$ (see \acronym{MIST} documentation; \citealt{WHO}), age $\tau$, metallicity $\feh$, extinction $A_V$, and distance $D$, which can be uniquely mapped to a point in the observed CMD;
The likelihood of a given set of these parameters, $\bs{\theta} = (\eep, \tau, \feh, A_V, D)$, is then computed from the photometry and (assumed Gaussian) photometric uncertainties given the predicted photometry.

@@

Ultimately we generate posterior samplings over the parameters $(\bs{\theta}, q)$ for each of the \Nisofit\ sources in the selected region of the \decam\ CMD using \texttt{PyMultinest} \citep{TODO}, and store the value of the prior evaluated at the location of each sample.
We then use these samples and interim prior values to construct our hierarchical model, as described below.
In the hierarchical model, we assume that the stellar parameters of each primary star in the selected region are either drawn from the cluster, or a background (stellar halo) population.
For the cluster, we assume the values are drawn from delta functions in age, metallicity, distance, and extinction, with the centroids of the delta functions $\bs{\alpha} = (\tau^*, \feh^*, D^*, A_V^*)$ as hyperparameters of the hierarchical inference.
For the background model, we assume the same priors as specified in \tablename~\ref{tbl:priors}.
The one additional parameter that must be included in this hierarchical model is the ``mixture weight'': the global fraction of sources that are likely cluster members, $f$.
To compute the likelihood of the hierarchical model, we use the individual posterior samplings to marginalize over the per-source $k$ stellar parameters $\bs{\theta}_k$ to compute the marginal likelihood $p(\bs{m}_k \given \bs{\alpha}, f)$, where $\bs{m}_k = (g, i)_k$ is the vector of photometric data;
This likelihood for a single source $k$ given a set of hyperparameters $(\bs{\alpha}, f)$ is
\begin{equation}
    p(\bs{m}_k \given \bs{\alpha}, f) = \int \dd \bs{\theta}_k \,
        p(\bs{m}_k \given \bs{\theta}_k) \,
        p(\bs{\theta}_k \given \bs{\alpha}, f) \quad .
\end{equation}
We employ the ``importance sampling trick'' (see, e.g., Appendix of \citealt{Price-Whelan:2018} or \citealt{Hogg:2008, Foreman-Mackey:2013} for other examples) to re-write an approximate form for this marginal likelihood as
\begin{equation}
    p(\bs{m}_k \given \bs{\alpha}, f) \approx \frac{\mathcal{Z}_k}{J} \,
        \sum_j^J \frac{p(\bs{\theta}_{kj} \given \bs{\alpha}, f)}{p(\bs{\theta}_{kj} \given \bs{\alpha}_0, f)}
        \label{eq:marglike}
\end{equation}
where the index $j$ specifies the index of one of $J$ posterior samples generated from the independent samplings (described above), $\mathcal{Z}_k$ is a constant, and the denominator, $p(\bs{\theta}_{kj} \given \bs{\alpha}_0, f)$, are the values of the interim prior used to do the independent samplings.
In this work, $K=\Nisofit$ and we adopt $J=2048$.

\todo{we ignore the $q$!}
With the marginal likelihood (\equationname~\ref{eq:marglike}), we then need to specify prior probability distributions for the hyperparameters $(\bs{\alpha}, f)$, and we can then generate posterior samples for the hyperparameters;
We use uniform priors for all of these, as summarized in \tablename~\ref{tbl:hyperpriors}.
We then use \texttt{emcee} \citep{Goodman:XX, Foreman-Mackey:2013} to sample from the posterior probability distribution for the hyperparameters given all of the photometric data,
\begin{equation}
    p(\bs{\alpha}, f \given \{\bs{m}_k\}) \propto
        p(\bs{\alpha}) \, p(f) \,
        \prod_k^K p(\bs{m}_k \given \bs{\alpha}, f) \quad .
\end{equation}
Here we use 64 walkers and run for an initial 128 steps to burn-in the sampler before running for a final 1024 steps.
\figurename~\ref{fig:hierarch-corner} shows a corner plot showing all 1D and 2D marginal posterior probability distributions estimated from the samples.

>>>>

We use astrometric data from the \gaia\ mission (\citealt{Prusti:2016}), data release 2 (\DR{2}; \citealt{Gaia-Collaboration:2018, Lindegren:2018}) to search for distant, comoving multiplets of blue stars.
Our original intent was to search for small, distant, comoving groups of blue horizontal branch stars to identify new candidate satellites of the Milky Way.
We therefore initially select all stars from \gaia\ with parallax $\varpi < 1$, color $-0.5 < (\bprp) < 0$, $G$-band magnitude $G < 20$, and Galactic latitude $|b| > 20^\circ$ (see Appendix~\ref{sec:queries} for database query).
We further exclude stars within a $15^\circ$ radius from the LMC, and a $8^\circ$ radius from the SMC --- 27,895 stars remain after these cuts.
We then cross-match this catalog to itself with both sky positions and proper motions: we search for pairs of stars that have separations $s < 0.5^\circ$ and proper motion differences $|\Delta \mu| < 0.5~\masyr$.
We then combine mutually-connected comoving pairs into small groups of stars that are colocated on the sky and comoving in proper motions, and remove groups that have $<4$ members.
We cross-match the mean sky positions of the groups to locations of local group galaxies \citep{McConnachie:2012} and Milky Way globular clusters \citep[2010 edition;][]{Harris:1996} and filter out all groups that lie within 1 degree of these known objects.
After these filters, just one group of blue, comoving stars remains at $(\textrm{RA}, \textrm{Dec}) \sim (179, -29)^\circ$.

@@

In detail, we start by using the \texttt{isochrones} package \citep{Morton:2015} to generate posterior samplings over stellar parameters for each individual source given its photometry and an interim prior.
We use the \acronym{MIST} \citep{Choi:2016} isochrone library, and \texttt{isochrones} automatically performs interpolation between the provided grid of stellar isochrones to evaluate predicted stellar parameters at a given location in the CMD.
The stellar parameters for each source are its ``equal evolutionary point'' number $\eep$ (see \acronym{MIST} documentation\footnote{\url{http://waps.cfa.harvard.edu/MIST/}}), age $\tau$, metallicity $\feh$, extinction $A_V$, and distance $D$, which can be uniquely mapped to a point in the observed CMD;
The likelihood of a given set of these parameters, $\bs{\theta} = (\eep, \tau, \feh, A_V, D)$, is then computed from the photometry and (assumed Gaussian) photometric uncertainties given the predicted photometry.

@@

Ultimately we generate posterior samplings over the parameters $(\bs{\theta}, q)$ for each of the \Nisofit\ sources in the selected region of the \decam\ CMD using \texttt{PyMultinest} \citep{TODO}, and store the value of the prior evaluated at the location of each sample.
We then use these samples and interim prior values to construct our hierarchical model, as described below.
In the hierarchical model, we assume that the stellar parameters of each primary star in the selected region are either drawn from the cluster, or a background (stellar halo) population.
For the cluster, we assume the values are drawn from delta functions in age, metallicity, distance, and extinction, with the centroids of the delta functions $\bs{\alpha} = (\tau^*, \feh^*, D^*, A_V^*)$ as hyperparameters of the hierarchical inference.
For the background model, we assume the same priors as specified in \tablename~\ref{tbl:priors}.
The one additional parameter that must be included in this hierarchical model is the ``mixture weight'': the global fraction of sources that are likely cluster members, $f$.
To compute the likelihood of the hierarchical model, we use the individual posterior samplings to marginalize over the per-source $k$ stellar parameters $\bs{\theta}_k$ to compute the marginal likelihood $p(\bs{m}_k \given \bs{\alpha}, f)$, where $\bs{m}_k = (g, i)_k$ is the vector of photometric data;
This likelihood for a single source $k$ given a set of hyperparameters $(\bs{\alpha}, f)$ is
\begin{equation}
    p(\bs{m}_k \given \bs{\alpha}, f) = \int \dd \bs{\theta}_k \,
        p(\bs{m}_k \given \bs{\theta}_k) \,
        p(\bs{\theta}_k \given \bs{\alpha}, f) \quad .
\end{equation}
We employ the ``importance sampling trick'' (see, e.g., Appendix of \citealt{Price-Whelan:2018} or \citealt{Hogg:2010, Foreman-Mackey:2014} for other examples) to re-write an approximate form for this marginal likelihood as
\begin{equation}
    p(\bs{m}_k \given \bs{\alpha}, f) \approx \frac{\mathcal{Z}_k}{J} \,
        \sum_j^J \frac{p(\bs{\theta}_{kj} \given \bs{\alpha}, f)}{p(\bs{\theta}_{kj} \given \bs{\alpha}_0, f)}
        \label{eq:marglike}
\end{equation}
where the index $j$ specifies the index of one of $J$ posterior samples generated from the independent samplings (described above), $\mathcal{Z}_k$ is a constant, and the denominator, $p(\bs{\theta}_{kj} \given \bs{\alpha}_0, f)$, are the values of the interim prior used to do the independent samplings.
In this work, $K=\Nisofit$ and we adopt $J=2048$.

\todo{we ignore the $q$!}
With the marginal likelihood (\equationname~\ref{eq:marglike}), we then need to specify prior probability distributions for the hyperparameters $(\bs{\alpha}, f)$, and we can then generate posterior samples for the hyperparameters;
We use uniform priors for all of these, as summarized in \tablename~\ref{tbl:hyperpriors}.
We then use \texttt{emcee} \citep{Goodman:2010, Foreman-Mackey:2013} to sample from the posterior probability distribution for the hyperparameters given all of the photometric data,
\begin{equation}
    p(\bs{\alpha}, f \given \{\bs{m}_k\}) \propto
        p(\bs{\alpha}) \, p(f) \,
        \prod_k^K p(\bs{m}_k \given \bs{\alpha}, f) \quad .
\end{equation}
Here we use 64 walkers and run for an initial 128 steps to burn-in the sampler before running for a final 1024 steps.
\figurename~\ref{fig:hierarch-corner} shows a corner plot showing all 1D and 2D marginal posterior probability distributions estimated from the samples.
